---
title: "assignment2_r_flights"
output: html_document
---

```{r}
library(tidyverse)
library(descr)
library(vcd)
```

```{r}
data <- read.csv('2008.csv.bz2')
head(data)
```

1. Find top-10 carriers in terms of the number of completed flights (UniqueCarrier column)?

Which of the listed below is not in your top-10 list?

— DL
— AA
— OO
— *EV*

```{r}
fl <- select(data, UniqueCarrier, Cancelled)
fl <- group_by(fl, UniqueCarrier)
fl <- summarise(fl, total=sum(Cancelled, na.rm=TRUE))
arrange(fl, desc(total))
```


2. Plot distributions of flight cancellation reasons (CancellationCode).

What is the most frequent reason for flight cancellation? (Use this link to translate codes into reasons)

— carrier
— *weather conditions*
— National Air System
— security reasons

```{r}
fl <- group_by(select(data, CancellationCode), CancellationCode)
fl <- count(fl)
fl <- filter(fl, CancellationCode != '')
reasons <- c('Carrier', 'Weather', 'National Air System', 'Security')
barplot(fl$n, names.arg = reasons, col = 'lightblue')
```


3. Which route is the most frequent, in terms of the number of flights?

(Take a look at 'Origin' and 'Dest' features. Consider A->B and B->A directions as different routes)

— New-York – Washington
— *San-Francisco – Los-Angeles*
— San-Jose – Dallas
— New-York – San-Francisco

```{r}
fl <- group_by(select(data, Origin, Dest), Origin, Dest)
arrange(count(fl), desc(n))
```


4. Find top-5 delayed routes (count how many times they were delayed on departure). From all flights on these 5 routes, count all flights with weather conditions contributing to a delay.

— 449
— 539
— 549
— *668*

```{r}
fl <- filter(select(data, Origin, Dest, DepDelay, WeatherDelay), DepDelay > 0)
fl <- group_by(fl, Origin, Dest)
fl_1 <- filter(fl, WeatherDelay > 0)
fl <- arrange(count(fl), desc(n))[1:5,]
origin = paste(fl$Origin[1:5])
dest = paste(fl$Dest[1:5])
weather_delay <- 0
for (i in (1:5)){
  weather_delay <- weather_delay + length(fl_1$WeatherDelay[fl_1$Origin == origin[i] & fl_1$Dest == dest[i]])
}
weather_delay
```



5. Examine the hourly distribution of departure times. For that, create a new series from DepTime, removing missing values.

Choose all correct statements:

— Flights are normally distributed within time interval [0-23] (Search for: Normal distribution, bell curve).
— Flights are uniformly distributed within time interval [0-23].
— *In the period from 0 am to 4 am there are considerably less flights than from 7 pm to 8 pm.*

```{r}
hours <- na.omit(data$DepTime)
hist(hours, col = 'pink')
```



6. Show how the number of flights changes through time (on the daily/weekly/monthly basis) and interpret the findings.

Choose all correct statements:

— *The number of flights during weekends is less than during weekdays (working days).*
— The lowest number of flights is on Sunday.
— *There are less flights during winter than during summer.*

```{r}
daily <- count(group_by(select(data, DayofMonth), DayofMonth))
barplot(daily$n, col = 'lightgreen', names.arg = daily$DayofMonth)
weekly <- count(group_by(select(data, DayOfWeek), DayOfWeek))
days = c('Mon', 'Tue', 'Wed', 'Thur', 'Fri', 'Sat', 'Sun')
barplot(weekly$n, col = 'violet', names.arg = days)
monthly <- count(group_by(select(data, Month), Month))
months = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sept', 'Oct', 'Nov', 'Dec')
barplot(monthly$n, col = 'lightyellow', names.arg = months)
```


7. Examine the distribution of cancellation reasons with time. Make a bar plot of cancellation reasons aggregated by months.

Choose all correct statements:

— *December has the highest rate of cancellations due to weather.*
— The highest rate of cancellations in September is due to Security reasons.
— *April's top cancellation reason is carriers.*
— Flights cancellations due to National Air System are more frequent than those due to carriers.

```{r}
cancelcode <- filter(data, data$CancellationCode != '')
args <- list(CancellationCode = c("", "Carr", "Wthr", "NAS", "Scrt"), Month = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sept', 'Oct', 'Nov', 'Dec'))
mosaic(data = cancelcode, CancellationCode ~ Month, set_labels = args, rot_labels = 0)
```


8. Which month has the greatest number of cancellations due to Carrier?

— May
— January
— September
— *April*

```{r}
cancelcode <- filter(data, data$CancellationCode != '')
args <- list(CancellationCode = c("", "Carr", "Wthr", "NAS", "Scrt"), Month = c('Jan', 'Feb', 'Mar', 'Apr', 'May', 'Jun', 'Jul', 'Aug', 'Sept', 'Oct', 'Nov', 'Dec'))
mosaic(data = cancelcode, CancellationCode ~ Month, set_labels = args, rot_labels = 0)
```



9. Identify the carrier with the greatest number of cancellations due to carrier in the corresponding month from the previous question.

— 9E
— EV
— HA
— *AA*

```{r}
fl <- filter(data, data$Month == 3 & data$Cancelled == 1)
fl <- count(group_by(select(fl, UniqueCarrier, Cancelled), UniqueCarrier))
arrange(fl, desc(n))
```



10. Examine median arrival and departure delays (in time) by carrier. Which carrier has the lowest median delay time for both arrivals and departures? Leave only non-negative values of delay times ('ArrDelay', 'DepDelay'). Boxplots can be helpful in this exercise, as well as it might be a good idea to remove outliers in order to build nice graphs. You can exclude delay time values higher than a corresponding .95 percentile.

— EV
— OO
— AA
— *AQ*

```{r}
fl <- filter(data, data$ArrDelay > 0, data$DepDelay > 0)
fl <- group_by(select(fl, UniqueCarrier, ArrDelay, DepDelay), UniqueCarrier)
arr <- summarise(fl, total=median(ArrDelay, na.rm = TRUE))
barplot(arr$total, names.arg = arr$UniqueCarrier, col = 'green')
dep <- summarise(fl, total=median(DepDelay, na.rm = TRUE))
barplot(dep$total, names.arg = dep$UniqueCarrier, col = 'yellow')
```