---
title: "assignment1_r_olympic"
output: html_document
---

```{r}
data <- read.csv('athlete_events.csv')
head(data)
```

1. How old were the youngest male and female participants of the 1996 Olympics?

— 16 and 15
— *14 and 12*
— 16 and 12
— 13 and 11

```{r}
min(data$Age[data$Sex == 'M' & data$Year == 1996], na.rm = TRUE)
min(data$Age[data$Sex == 'F' & data$Year == 1996], na.rm = TRUE)
```

2. What was the percentage of male gymnasts among all the male participants of the 2000 Olympics? Round the answer to the first decimal.

Hint: here and further if needed drop duplicated sportsmen to count only unique ones.

— 0.2
— *1.5*
— 2.5
— 7.7

```{r}
male_part <- length(unique(data$ID[data$Sex == 'M' & data$Year == 2000]))
male_gym <- length(unique(data$ID[data$Sex == 'M' & data$Year == 2000 & data$Sport == 'Gymnastics']))
round(male_gym / male_part * 100, 1)
```


3. What are the mean and standard deviation of height for female basketball players participated in the 2000 Olympics? Round the answer to the first decimal.

— 178.5 and 7.2
— 179.4 and 10
— 180.7 and 6.7
— *182.4 and 9.1*

```{r}
round(mean(data$Height[data$Sex == 'F' & data$Sport == 'Basketball' & data$Year == 2000], na.rm = TRUE), 1)
round(sd(data$Height[data$Sex == 'F' & data$Sport == 'Basketball' & data$Year == 2000], na.rm = TRUE), 1)
```

4. Find a sportsperson participated in the 2002 Olympics, with the highest weight among other participants of the same Olympics. What sport did he or she do?

— Judo
— *Bobsleigh*
— Weightlifting
— Boxing

```{r}
max_weight <- max(data$Weight[data$Year == 2002], na.rm = TRUE)
sport <- data$Sport[data$Year == 2002 & data$Weight == max_weight]
sport[!is.na(sport)]
```

5. How many times did Pawe Abratkiewicz participate in the Olympics held in different years?

— 0
— 1
— 2
— *3*

```{r}
length(unique(data$Year[data$Name == 'Pawe Abratkiewicz']))
```

6. How many silver medals in tennis did Australia win at the 2000 Olympics?

— 0
— 1
— *2*
— 3

```{r}
n <- data$ID[data$Medal == 'Silver' & data$Sport == 'Tennis' & data$Team == 'Australia' & data$Year == 2000]
length(unique(!is.na(n)))
```

7. Is it true that Switzerland won fewer medals than Serbia at the 2016 Olympics? Do not consider NaN values in Medal column.

— *Yes*
— No

```{r}
Swtz <- data$Medal[data$Team == 'Switzerland' & data$Year == 2016]
Swtz <- length(Swtz[!is.na(Swtz)])
Srb <- data$Medal[data$Team == 'Serbia' & data$Year == 2016]
Srb <- length(Srb[!is.na(Srb)])
if (Swtz < Srb) print('Yes') else print('No')
```

8. What age category did the fewest and the most participants of the 2014 Olympics belong to?

— *[45-55] and [25-35) correspondingly*
— [45-55] and [15-25) correspondingly
— [35-45] and [25-35) correspondingly
— [45-55] and [35-45) correspondingly

```{r}
age_35 = length(unique(data$ID[data$Age >= 35 & data$Age <= 45 & data$Year == 2014]))
age_45 = length(unique(data$ID[data$Age >= 45 & data$Age <= 55 & data$Year == 2014]))
if (min(age_35, age_45) == age_35){
  print('[35-45]')
} else {
  print('[45-55]')
}
age_35 = length(unique(data$ID[data$Age >= 35 & data$Age < 45 & data$Year == 2014]))
age_25 = length(unique(data$ID[data$Age >= 25 & data$Age < 35 & data$Year == 2014]))
age_15 = length(unique(data$ID[data$Age >= 15 & data$Age < 25 & data$Year == 2014]))
if (max(age_15, age_25, age_35) == age_35){
  print('[35-45)')
} else if (max(age_15, age_25, age_35) == age_25){
  print('[25-35)')
} else{
  print('[15-25)')
}
```

9. Is it true that there were Summer Olympics held in Lake Placid? Is it true that there were Winter Olympics held in Sankt Moritz?

— Yes, Yes
— Yes, No
— *No, Yes*
— No, No

```{r}
if (length(data$ID[data$Season == 'Summer' & data$City == 'Lake Placid']) == 0){
  print('No')
} else {
  print('Yes')
}
if (length(data$ID[data$Season == 'Winter' & data$City == 'Sankt Moritz']) == 0){
  print('No')
} else {
  print('Yes')
}
```

10. What is the absolute difference between the number of unique sports at the 1995 Olympics and 2016 Olympics?

— 16
— 24
— 26
— *34*

```{r}
abs(length(unique(data$Sport[data$Year == 1995])) - length(unique(data$Sport[data$Year == 2016])))
```